\documentclass[letterpaper, 12pt]{report}
\usepackage[margin=0.75in]{geometry}
%\usepackage[english]{babel} 
\usepackage[T1]{fontenc}
\usepackage{graphicx, lipsum, textcomp, float} %figure formatting
\usepackage{hyperref} %referancing package
%\usepackage[version=4]{mhchem}  %chemical equations and formulas
\usepackage{chemformula}
%\usepackage{advdate, datenumber} %date packages
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{multirow}
%\usepackage{comment}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{adjustbox}
\usepackage{array, booktabs}
\usepackage{nameref}
\usepackage{etaremune}
\DeclareUnicodeCharacter{2212}{-}
%Listings
\usepackage{minted}
%\usepackage[finalizecache,cachedir=.]{minted}
%\usepackage[frozencache,cachedir=.]{minted}

%Font
\renewcommand{\thelinenumber}{\raisebox{1pt}{\textcolor[RGB]{200,200,200}{\arabic{linenumber}}}}
%\usepackage[scaled]{beramono}
%\usepackage{tgpagella}
%\renewcommand{\familydefault}{\sfdefault} 
\definecolor{darkgreen}{rgb}{0.05, 0.3, 0.1}

%\usepackage[htt]{hyphenat} %texttt hyphenation breaks

\let\oldtexttt\texttt
\renewcommand{\texttt}[1]{\oldtexttt{\textcolor{darkgreen}{#1}}}


% Line spacing (double for title page, single for TOC, and 1.5 for body)
\usepackage{setspace}

% Bibliography
\usepackage[
    style=ieee, 
    isbn=false, 
    url=true, 
    natbib=true, 
    backend=bibtex,
    maxcitenames=1,
    mincitenames=1
    ]{biblatex}
\addbibresource{references.bib}
\renewcommand*{\bibfont}{\footnotesize}

% Smaller figure captions 
\usepackage{caption}
\captionsetup[figure]{font=normalsize, labelfont=normalsize}

% Section titles
\usepackage{titlesec}
\setcounter{secnumdepth}{4}

% Chapter
\titleformat{\chapter}[display]
{\Large\bfseries\centering}
{Chapter \thechapter}{0.5em}{}[\vspace{2ex}\titlerule]
\titlespacing*{\chapter}{0pt}{0pt}{30pt}

% Section
\titleformat{\section}[hang]
{\large\bfseries}
{\thesection}{0.5em}{}

% Subsection
\titleformat{\subsection}[hang]
{\large\bfseries}
{\thesubsection}{0.5em}{}

% Subsubection
\titleformat{\subsubsection}[hang]
{\large\bfseries}
{\thesubsubsection}{0.5em}{}

% less section skip in the table of contents
\usepackage{tocbasic}

% Table
\renewcommand{\arraystretch}{1.2}

% Section
\DeclareTOCStyleEntry[
  beforeskip=.2em plus 1pt,% default is 1em plus 1pt
  pagenumberformat=\textbf
]{tocline}{section}

% Chapter
\DeclareTOCStyleEntry[
  entrynumberformat=\entrywithprefix{\chaptername},
  dynnumwidth
]{tocline}{chapter}
\newcommand*\entrywithprefix[2]{#1~#2}

% Appendix
\usepackage[toc]{appendix}
\newcommand{\mypart}[1]{\thispagestyle{empty}\part*{#1}}%\addtocounter{page}{-1}}

% MISC
\setlength\parindent{6pt} %paragraph indentation
\setlength{\parskip}{6pt} %paragraph spacing

%set hyperlinks colors
\definecolor{mypurple}{RGB}{140,54,140}
\definecolor{homered}{RGB}{127, 0, 10}
\definecolor{officeorange}{RGB}{204, 75, 0}
\definecolor{mauroblue}{RGB}{53, 48, 217}
\definecolor{citegreen}{RGB}{15, 133, 13}
\definecolor{hyperlinkpurple}{RGB}{42, 0, 163}
\definecolor{subtlegray}{gray}{0.98}
\definecolor{subduedgray}{gray}{0.75}
\hypersetup{
    colorlinks=true,
    linkcolor=hyperlinkpurple,
    filecolor=mypurple,      
    urlcolor=teal,
    citecolor=citegreen
}

% Macros:
% Full number and reference name hyperlinking
\newcommand*{\fullref}[1]{\hyperref[{#1}]{\ref*{#1} on \nameref*{#1}}}
% Acknoledgments on per-chapter basis
\newcommand{\acknowledge}[1]{\textit{
\small
Acknowledgment: #1
}}
% TODO markers.
\newcommand{\todo}{
\begin{center}
\textcolor{mauroblue}{
\textit{
This section is currently under preparation.
}}
\end{center}
}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   DOCUMENT   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

% Front matter manually formatted according to the rules 
\pagenumbering{roman}
\thispagestyle{empty}
\setstretch{1}

% Title page body
{
\centering
The Pennsylvania State University\\
The Graduate School\\
\vfill
\setstretch{2}
{
\fontsize{14}{16}\selectfont
\textbf{INVESTIGATION OF ATOMIC ENVIRONMENTS FROM COMPUTATIONAL THERMODYNAMICS: APPLICATIONS IN INTERMETALLIC CATALYSTS AND MOLTEN SALTS}\\
}
\vfill
A Dissertation in\\
Materials Science and Engineering\\
by\\
Rushi Gong\\
\vfill
© 2024 Rushi Gong\\
\setstretch{1}
\vfill
Submitted in Partial Fulfillment\\
of the Requirements\\
for the Degree of\\

\vfill
Doctor of Philosophy\\
\vfill
December 2024\\
\vfill
}

% Committee page
\newpage
\setstretch{1.5}
\setlength\parindent{0pt} %no paragraph indentation

The dissertation of Rushi Gong was reviewed and approved by the following:\\

\textbf{Zi-Kui Liu}\\
Dorothy Pate Enright Professor at the Department of Materials Science and Engineering\\
Dissertation Advisor\\
Chair of the Committee\\

\textbf{Michael Janik}\\
Professor of Chemical Engineering\\

\textbf{Hojong Kim}\\
Associate Professor of Materials Science and Engineering\\

\textbf{John Mauro}\\
Dorothy Pate Enright Professor and Associate Head for Graduate Education\\
Chair, Intercollege Graduate Degree Program in Materials Science and Engineering\\
Program Head

\vfill

\newpage
\chapter*{Abstract}
Understanding atomic environments is essential for optimizing the performance and properties of materials by providing insights into structure-property relationships. CALPHAD modeling, which describes the Gibbs energies of individual phases, is a key tool for describing multi-component thermodynamics. For complex multi-component solution phases, advanced thermodynamic models are required to capture the inherent complexities of atomic environments in the phase, such as short-range ordering. A significant challenge in CALPHAD modeling is comparing various thermodynamic models and selecting the most appropriate one for each phase. Open-source software tools, PyCalphad and ESPEI, have enabled high-throughput CALPHAD modeling with uncertainty quantification. The integration of Bayesian parameter estimation within the Markov Chain Monte Carlo approach in ESPEI facilitates Bayesian model selection in CALPHAD modeling. 

This dissertation will discuss the selection and statistical comparison of thermodynamic models for describing atomic environments in materials, focusing on intermetallic catalysts and molten salts. Employing a four-sublattice model for the Pd-Zn-based $\gamma$-brass phase successfully predicts the site occupancy of active metals within intermetallic structures. This atomic control of active-site ensembles in Pd-Zn-based $\gamma$-brass intermetallics allows for the precise tailoring of activity and selectivity in hydrogenation reactions. For complex molten salts, candidate models such as the associate model, two-sublattice ionic model, and modified quasichemical model with quadruplets approximation are considered to describe short-range ordering. Bayesian statistics are applied to compare these models and identify the most favorable one for molten salt liquids, further enhancing the accuracy of predictive modeling of critical characteristics. Additionally, a template generator has been developed to allow users to customize thermodynamic models within PyCalphad. These advancements provide the community with extensive opportunities to comprehensively evaluate thermodynamic modeling with uncertainty quantification and model selection, accelerating the optimization process in data-driven materials design and discovery.

\setstretch{1}
\newpage
\setcounter{tocdepth}{3}
\tableofcontents

\newpage
\renewcommand{\listfigurename}{List of Figures}
\addcontentsline{toc}{chapter}{\listfigurename}
\listoffigures


\newpage
\renewcommand{\listtablename}{List of Tables}
\addcontentsline{toc}{chapter}{\listtablename}
\listoftables

\newpage
\chapter*{Acknowledgments}
\label{acknowledgments}
\addcontentsline{toc}{chapter}{\nameref{acknowledgments}}

I would like to express heartfelt thanks to my advisor, Dr. Zi-Kui Liu, for guiding me over the last five years. His passion for research has deeply influenced me. I will always be grateful for his continuous guidance and encouragement in my professional and personal development.

I would like to thank the current and former colleagues from Phases Research Lab, Dr. Shun-Li Shang, Dr. Yi Wang, Dr. Brandon Bocklund, Dr. Jorge Paz Soldan Palma, Dr. John Shimanek, Dr. Hui Sun, Dr. Adam Krajewski, Dr. Nigel Hew, Shuang Lin, Alexander Richter, Zhening Yang, Luke Myers, and Ricardo Amaral. I am so fortunate to have their guidance and have this opportunity to grow with them.

I would like to thank Dr. Michael Janik for supporting the intermetallics work and guiding my early development during my PhD journey. I am also grateful for the opportunity to complete two internships at Argonne National Laboratory. I would like to thank my advisor at Argonne National Laboratory Nuclear Science and Engineering Division, Dr. Shayan Shahbazi, who prompted me to deepen my understanding of molten salts. I would like to thank the support from Dr. Hojong Kim and Dr. Xiaofeng Guo for collaborations on molten salts projects.

I would like to thank my friends at State College and all the warm people I met at SCCAC. Thanks to my boyfriend Jiayang Wang for coming on this journey with me. I am so grateful to have Ori in my life, who lights up my world. Finally, I would like to express my deepest thanks to my family: my Mom, Mengjun Hu, and my Dad, Wei Gong, whose unwavering support will always be my greatest strength. I truly would not be where I am today without my parents.

This work was made possible by the financial support and training provided by the US Department of Energy (DOE) Office of Science, Office of Basic Energy Sciences, Catalysis Division via Award No. DE-SC0020147, DOE Nuclear Energy University Program (NEUP) via Award Nos. DE-NE0008945 and DE-NE0009288, and the Nuclear Energy Advanced Modeling and Simulation (NEAMS) program. Simulations were performed on the Roar supercomputer at the Pennsylvania State University's Institute for Computational and Data Sciences (ICDS) and the ACCESS supported by the National Science Foundation (NSF) with Grant No. ACI-1548562. Any opinions, findings, conclusions, or recommendations expressed in this publication are those of the author and do not necessarily reflect the views of the funding agencies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\setlength\parindent{2em} %paragraph indentation
\setstretch{1.5}
\pagenumbering{arabic}

\include{introduction}

\include{methodology}

\include{intermetallics}

\include{moltensalts}

\include{models}

\include{conclusion}

\printbibliography[
heading=bibintoc,
title={Bibliography}
]

\newpage
\chapter*{Vita}
\thispagestyle{empty}
\vspace{-6pt}

{\small
Rushi Gong was born on October 15, 1997, in Yiyang, Hunan Province, China. She grew up in Changsha, Hunan, and later moved to Shenzhen, Guangdong. In 2019, she graduated from Beihang University in Beijing, China, with a B.S. in Materials Science and Engineering and a minor in Mathematics. In 2020, Rushi began her Ph.D. studies at the Pennsylvania State University, advised by Dr. Zi-Kui Liu. Listed below are the publications she contributed to during her Ph.D.:
\fontsize{10}{10}\selectfont
\input{publications}
}

\end{document}
